Behavior Driven Development in Haskell

### 2012-06-02 paper.tex
Finalize terminology

Unfortunately, the type system of Haskell does not always guarantee
correct semantics.
For example, it does not ensure that division by zero does not occur.
An other example is, that instances of {\tt Monoid} \emph{should} be
monoids in the mathematical sense, but the type system does not
enforce this.  It is the programmers responsibility to ensure that
instance definitions \emph{behave} as \emph{expected}.  Hence
automated testing is useful.

However, some widely used Haskell libraries do not have automated
tests.  For example, the Haskell Platform 2011.4.0.0 contains 23
packages, not counting the 19 core libraries that come with GHC.  Only
8 of them have a test suite, 5 provide example code, and 10 have
neither tests nor examples.  A lack of tests is not only a quality and
maintainability issue.  It also is a lack of documentation.  Tests
document how code is supposed to be used, and tests document the
behavior of code.

@@

Hspec is thus useful for \emph{code documentation}, but we observed that parts
of a specification can be useful as user documentation. To that end,
Python's {\tt doctest} module~\cite{doctest} is of interest for us.
It is not commonly described as a BDD tool, but we think it can be
used for BDD.
There is a Haskell version called Doctest\footnote{%
http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
the standard documentation mechanism for Haskell.
It extracts \emph{interactive examples} from Haddock comments and
verifies them automatically.

@@

\item
    We introduce a set of combinators to express expectations about
    code examples for use with Hspec
    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb})

@@

\item[spec]
    A spec is similar to a test suite in TDD.  But it is written with
    specification terminology, and produces a specification document
    when run.
\end{description}

\begin{quote}
    \emph TBD: Check, terminology for consistency, and remove
    unnecessary terminology (if any).  This has to be done as one of
    the last steps.
\end{quote}

\begin{description}
\item[developer]
    A programmer who develops a library.

\item[user]
    A programmer who uses a library.

\item[user documentation]
    Explanation of functions for users.
\item[code documentation]
    Explanation of functions for developers.

\item[non-quantified property]
    Property which does not take parameters.

\item[quantified property]
    Property which takes parameters.

@@

\noindent Hspec does not yet provide a mechanism to express
\emph{expectations} with that terminology.  The developer has to use
HUnit assertions instead.

@@

We are grateful to David Waern, Haddock's maintainer, for reviewing and
accepting our patches to Haddock.
We would like to thank
Nick Smallbone for having improved QuickCheck so that
non-quantified properties are tested just once.

>>>>

Unfortunately, the type system of Haskell does not always guarantee
correct semantics.
For example, it does not ensure that division by zero does not occur.
Another example is, that instances of {\tt Monoid} should be
monoids in the mathematical sense, but the type system does not
enforce this.  It is the programmers responsibility to ensure that
instance definitions behave as expected.  Hence
automated testing is useful.

However, some widely used Haskell libraries do not have automated
tests.  For example, the Haskell Platform 2011.4.0.0 contains 23
packages, not counting the 19 core libraries that come with GHC.  Only
8 of them have a test suite, 5 provide example code, and 10 have
neither tests nor examples.  A lack of tests is not only a quality and
maintainability issue.  It also is a lack of documentation.  Tests
document how code is supposed to be used, and tests document the
\emph{behavior} of code.

@@

Hspec is thus useful for \emph{code documentation}, but we observed that parts
of a specification can be useful as \emph{user documentation}. To that end,
Python's {\tt doctest} module~\cite{doctest} is of interest for us.
It is not commonly described as a BDD tool, but we think it can be
used for BDD.
There is a Haskell version called Doctest\footnote{%
http://hackage.haskell.org/package/doctest}, which is integrated with Haddock~\cite{haddock},
the standard documentation mechanism for Haskell.
It extracts \emph{interactive examples} from Haddock comments and
verifies them automatically.

@@

\item
    We introduce a set of combinators to express \emph{expectations} about
    code examples for use with Hspec
    (Section~\ref{sec:code-doc} and Section~\ref{sec:hspec-comb})

@@

\item[spec]
    A spec is similar to a test suite in TDD.  But it is written with
    specification terminology, and produces a specification document
    when run.

\item[user documentation]
    User documentation means the documentation for users of a library.

\item[code documentation]
    Code documentation means a detailed description, or specification,
    of a library.

@@

\noindent Hspec does not yet provide a mechanism to express
expectations with that terminology.  The developer has to use
HUnit assertions instead.

@@

We are grateful to David Waern, Haddock's maintainer, for reviewing
and accepting our patches to Haddock.
%We would like to thank
%Nick Smallbone for having improved QuickCheck so that
%non-quantified properties are tested just once.
